---
title: "Practical Machine Learning Course Project"
author: "Ksenia Slivkina"
date: "23 09 2020"
output: html_document
---

Course Project.

I possess data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. In this project i used to predict the manner in which these people did the exercise.

Plan of the work:
* load the data.
* implement cross-validation.
* clean the data.
* built a model with Random Forests alforithm.
* apply it for testing set.
* predict 20 different test cases using this model.

Loading all needed libraries.
```{r}
library(caret)
library(randomForest)
library(dplyr)
```
Reading the data for training and testing.
```{r}
Datatraining = read.csv("pml-training.csv")
Datatesting = read.csv("pml-testing.csv")
```
Using cross-validation to creat dataset "train" ,and devide it into 60% of training set (for model building) and 40% of testing set (for tesing model).
```{r}
set.seed(1234)
train = createDataPartition(y = Datatraining$classe,p = .60,list = FALSE)
training = Datatraining[train,]
testing = Datatraining[-train,]
```
Aplly "nearZeroVar" function to find variables that have no variability and won't be useful covariates, then remove them from training set.
```{r}
training = training %>%
  select_at(vars(-one_of(nearZeroVar(., names = TRUE))))

```
Continue cleaning the training set: 
1) delete colums that are useless for making predictions:ID variable, user_name, all varianles with timestamps and window data.
2) keep colums with less than or equal to 95% of NA’s.
```{r}
DELCOL = grep("X|user_name|timestamp|window", colnames(training), value = FALSE) 
trainingclean = training[,-DELCOL]
# replace missing data by NA's
trainingclean[trainingclean == ""] = NA
NAcount = apply(trainingclean, 2, function(k) sum(is.na(k)))/nrow(trainingclean)
trainingclean = trainingclean[NAcount <= 0.95]
```
I use Random Rorests alforithm to create a model as it is one of the most highly accurate methods for prediction.
```{r}
modFit = randomForest(trainingclean$classe ~ ., data = trainingclean,do.trace = FALSE)
modFit 
```
Error rate is only 0.61% that seems to be good.

Aplly cleaning algorithm to the testing set.
```{r}
testing = testing %>%
  select_at(vars(-one_of(nearZeroVar(., names = TRUE))))
testingclean = testing[,-DELCOL]
testingclean[testingclean == ""] <- NA
NAcount = apply(testingclean, 2, function(k) sum(is.na(k)))/nrow(testingclean)
testingclean = testingclean[NAcount <= 0.95]
```
Make prediction and create confusion matrix to test results.
```{r}
prediction_testingCl <- predict(modFit, testingclean, type = "class")
confusionMatrix(testingclean$classe,prediction_testingCl)
```
Accuracy on the testing set is 99,4 % , this is very high result.

Check on the test set and look at predictions for 20 test cases.
```{r}
Datatesting = Datatesting %>%
  select_at(vars(-one_of(nearZeroVar(., names = TRUE))))
testdataclean = Datatesting[,-DELCOL]
testdataclean[testdataclean == ""] <- NA
NAcount = apply(testdataclean, 2, function(k) sum(is.na(k)))/nrow(testdataclean)
testdataclean = testdataclean[NAcount <= 0.95]
testdataclean$classe = predict(modFit,testdataclean)
```

```{r}
testdataclean$classe
````